---
title: "Demand Forecasting"
output: html_notebook
---

```{r}
library(tidyverse)
library(zoo)
library(lubridate)
```


## Exploratory data analysis

```{r}
train_data = read.csv("train.csv" , stringsAsFactors = F)
head(train_data)
```

No NA values - the heavens have ssmiled upon us - it is quite rare to find data like this.  However the store and item numbers are being read as numeric variable instead of factor variables - we need to fix that.
```{r}
train_data$store = as.character(train_data$store)
train_data$item = as.character(train_data$item)
train_data$date = as.Date(train_data$date)
table(train_data[c("store", "item")])

```

Again very consistent data - all stores have observations for all items 

```{r}
train_data %>%
  mutate(month = as.yearmon(date)) %>%
  group_by(month) %>%
  summarise( total_sales = sum(sales) ) %>%
  ggplot() +geom_line(aes(x = month , y = total_sales)) + ggtitle("Monthly Sales - Total")
```
Two components
```{r}
train_data %>%
  mutate(year =  as.numeric(format.Date( train_data$date , format = "%Y")) ) %>%
  group_by(year) %>%
  summarise( total_sales = sum(sales) ) %>%
  ggplot() +geom_line(aes(x = year , y = total_sales)) + ggtitle("Yearly Sales - Total")
```
Trend in sales yearly - Sales increasing every year
```{r}
train_data %>%
  group_by(date) %>%
  summarize(total_daily_sales = sum(sales)) %>%
  mutate(week =  week(date) , year = year(date) ,day =  weekdays(date) ) %>%
  group_by(week, year) %>%
  mutate( normalized_sales = total_daily_sales/ sum(total_daily_sales) ) %>%
  group_by(day) %>%
  summarize(avg_normalied_sales = mean(normalized_sales , na.rm = T)) %>%
  mutate(day = factor(day , levels = c("Monday" , "Tuesday","Wednesday", "Thursday" , "Friday", "Saturday" , "Sunday"))) %>%
  ggplot() +geom_bar(aes(x = day , y = avg_normalied_sales) , stat = "identity") + ggtitle("Yearly Sales - Total") +
  scale_y_continuous(labels = scales::percent)
```




```{r}
train_data %>%
  group_by(date) %>%
  summarize(total_daily_sales = sum(sales)) %>%
  mutate(month =  month(date) , year = year(date) ,day =  weekdays(date) ) %>%
  group_by(month, year) %>%
  mutate( normalized_sales = total_daily_sales/ sum(total_daily_sales) ) %>%
  group_by(day) %>%
  summarize(avg_normalied_sales = mean(normalized_sales , na.rm = T)) %>%
  mutate(day = factor(day , levels = c("Monday" , "Tuesday","Wednesday", "Thursday" , "Friday", "Saturday" , "Sunday"))) %>%
  ggplot() +geom_bar(aes(x = day , y = avg_normalied_sales) , stat = "identity") + ggtitle("Weekly Trend") +
  scale_y_continuous(labels = scales::percent)
```
A weekly effectt is apparant where sales increase as the week goes by - maxiumum on weekends.

```{r}
total_sales_by_week = train_data %>%
  mutate(week =  week(date) , year = year(date)) %>%
  group_by(year,week) %>%
  summarize( total_sales =sum(sales))

acf(total_sales_by_week$total_sales, lag.max = 104)
```

```{r}
total_sales_by_month = train_data %>%
  mutate(month =  month(date) , year = year(date)) %>%
  group_by(year,month) %>%
  summarize( total_sales =sum(sales))

acf(total_sales_by_month$total_sales)
```

Yearly seasonality is apprant in the total sales

formula = (beta_yer* year + alpha_year)()

log(sales) = (beta_yer* year + alpha_year)